---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# litllm

<!-- badges: start -->

<!-- badges: end -->

The goal of `litllm` is to help you mine pdfs for scholarly analysis.

## Installation

You can install the development version of litllm from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("avahoffman/litllm")
```

## Getting started

```{r eval=F}
library(litllm)
```

```{r echo=F, message=F}
devtools::load_all()
```

`litllm` uses the OpenAI API. You'll need an API key and internet connection to use it. Check that you have an API key saved in your environment:

```{r}
ll_check_api_key()
```

_Note: The OpenAI API costs money. It is very cheap (around $0.80 for analyzing 100 pdfs), but you'll need to add credits through their interface._

Create a subdirectory in your working directory. We typically call ours "pdf/". Add the pdf files for any publications you'd like to analyze to this directory. 

_Note: We recommend starting with ~2 files so you can make sure everything is working as expected._

## Getting authors and titles

The first thing you might want to do is iterate through all your pdf files, finding the article's title and author information:

```{r eval=F}
# look at files in the "pdf/" directory
author_df <- ll_extract_authors("pdf/")
```

`ll_extract_authors()` produces a tibble with a column "paper_id" (the file name), "title" (the article title), and "auth" (individual author information). Each author will be on a separate row.
