---
title: "litllm Vignette"
format: html
editor: source
---

## `litllm` Vignette

### Getting Set Up

Install `litllm`.

```{r}
remotes::install_github(...)
```

Load `litllm` and `dplyr`.

```{r}
library(litllm)
library(dplyr)
```

First, you'll need an OpenAI API key. Check that you have one.

```{r}
ll_check_api_key()
```

Download this file and ensure it's in your working directory.

```{r}
download.file(
  url = "https://www.annualreviews.org/docserver/fulltext/genom/25/1/annurev-genom-021623-081639.pdf?expires=1731900753&id=id&accname=guest&checksum=AC39C6258EF7E19E7426CB44F2B37958",
  destfile = "Taylor_etal.pdf"
  )
list.files(pattern = "pdf")
```

```         
[1]"Taylor_etal.pdf"
```

### Mining Authors and Institutions

```{r}
authors <- ll_extract_authors("Taylor_etal.pdf")
glimpse(authors)
```

```         
Rows: 11
Columns: 3
$ paper_id <chr> "Taylor_etal.pdf", "Taylor_etal.pdf", "Taylor_etal.pdf", "Ta…
$ title    <chr> "Beyond the Human Genome Project: The Age of Complete Human …
$ auth     <chr> "Dylan J. Taylor, Department of Biology, Johns Hopkins Unive…
```

You can view the data in raw format as well.

```{r}
authors_raw <- ll_extract_authors("Taylor_etal.pdf", clean_authors = F)
authors_raw
```

```
$Taylor_etal.pdf
[1] "Title: Beyond the Human Genome Project: The Age of Complete Human Genome Sequences and Pangenome References\n\nAuthor(s):\nDylan J. Taylor, Department of Biology, Johns Hopkins University, Baltimore, Maryland, USA\nJordan M. Eizenga, Genomics Institute, University of California, Santa Cruz...
```

You can clean and pull institutions from each author+affiliation entry.

```{r}
institutions <- ll_pull_institutions(authors$auth)
authors$inst <- institutions
glimpse(authors)
```

```
Rows: 11
Columns: 4
$ paper_id <chr> "Taylor_etal.pdf", "Taylor_etal.pdf", "Tayl…
$ title    <chr> "Beyond the Human Genome Project: The Age o…
$ auth     <chr> "Dylan J. Taylor, Department of Biology, Jo…
$ inst     <chr> "Johns Hopkins University", "University of …
```

### Getting Institution Information

We might be interested in information about the institutions. But first let's limit it to unique institutions.

```{r}
distinct_institutions <- unique(authors$inst)
```

```
[1] "Johns Hopkins University"                                                                 
[2] "University of California, Santa Cruz"                                                     
[3] "Johns Hopkins University School of Medicine"                                              
[4] "Icahn School of Medicine at Mount Sinai"                                                  
[5] "Institut de Recherche en Santé Digestive, Université de Toulouse, INSERM, INRA, ENVT, UPS"
```

Let's see if these institutions have a large student body.

```{r}
is_large <- 
  ll_check_institution_type(
    distinct_institutions, 
    logical_check = "an institution with >5000 students"
    )
is_large
```

```
# A tibble: 5 × 2
  name                                          an_institution_with_…¹
  <chr>                                         <chr>                 
1 Johns Hopkins University                      TRUE                  
2 University of California, Santa Cruz          TRUE                  
3 Johns Hopkins University School of Medicine   FALSE                 
4 Icahn School of Medicine at Mount Sinai       FALSE                 
5 Institut de Recherche en Santé Digestive, Un… FALSE                 
# ℹ abbreviated name: ¹​an_institution_with_5000_students
```

Join to `authors` to build your dataset!

```{r}
authors <- authors %>% left_join(is_large, by = c("inst" = "name"))
```

```
Rows: 11
Columns: 5
$ paper_id                          <chr> "Taylor_etal.pdf", "Taylor_etal.pd…
$ title                             <chr> "Beyond the Human Genome Project: …
$ auth                              <chr> "Dylan J. Taylor, Department of Bi…
$ inst                              <chr> "Johns Hopkins University", "Unive…
$ an_institution_with_5000_students <chr> "TRUE", "TRUE", "TRUE", "TRUE", "F…
```
